### 2015-11-09 chapter5.tex

\chapter{Conclusion}

Many artworks are introduced. They have differences in terms of purpose and approached to the trash. Some of them only valued from its plastics effects. Some of them turned to the performative act.

\textbf{Future Works.} Web site will expand in time. New notebooks from different papers (in terms of location, meaning, usage purpose)

There is also discard is being generated during the production process of papers. as an production process how to use them? need to find a way because varda uses this discard items. etc. even if work is generated from discard items. again always new discard items are generated.

% Bu bölümde gene toplu bir özet yapılacak. Teker teker argumanlardan bahsedilecek. ve bu argumanlerın yeri belirtilecek.

% O zaman yarın gene tüm yazdıklarımı okuyorum, nasıl birleşeceğine karar veriyorum.