### 2015-11-11 chapter5.tex

\chapter{Conclusion}

%****************************************
% Savunma.
% Yani aslında bu tezdeki fikirler, işler birer supportive yaklaşımlar. neler yapılmış onları listesi. literatür taraması. ama tabi benim işime yarayanlar. 
%........................................

\epigraph{If I seem to be over-interested in junk, it is because I am, and I have a lot of it, too --- half a garage full of bits and broken pieces. I use these things for repairing other things\ldots But it can be seen that I do have a genuine and almost miserly interest in worthless objects. My excuse is that in this era of planned obsolescence, when a thing breaks I can usually find something in my collection to repair it --- a toilet, or a motor, or a lawn mower. But I guess the truth is that I simply like junk.}{\hfill---John Steinbeck, \textit{Travels with Charley, 1962}}

Many artworks are introduced. They have differences in terms of purpose and approached to the trash. Some of them only valued from its plastics effects. Some of them turned to the performative act.

\textbf{Future Works.} Web site will expand in time. New notebooks from different papers (in terms of location, meaning, usage purpose)

There is also discard is being generated during the production process of papers. as an production process how to use them? need to find a way because varda uses this discard items. etc. even if work is generated from discard items. again always new discard items are generated.