### 2015-12-16 main.tex
Thesis format edits
Former-commit-id: d1dea8cb723a1b15083a10b05c3c9849cab174b9

\documentclass[12pt]{report}
\usepackage[hidelinks]{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[turkish,english,shorthands=:!]{babel}
\usepackage{apacite}
\usepackage[a4paper,left=40mm,right=25mm,top=25mm,bottom=25mm]{geometry}
\usepackage{epigraph}
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[usenames,dvipsnames]{color}
\usepackage[titletoc]{appendix}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{afterpage}
\usepackage[doublespacing]{setspace}
\usepackage{tocloft}

\usepackage{chngcntr}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}

\usepackage{natbib}
\setlength{\bibsep}{0.0pt}

% Adding extra space between paragraphs
\setlength{\parskip}{\baselineskip}

% To align left todos
\reversemarginpar

% Chapter titles
\usepackage{titlesec}
\titleformat{\chapter}[display]
        {\normalfont\fontsize{14pt}{12pt}\centering\bfseries}{\MakeUppercase{\chaptertitlename}\  \thechapter}{2\baselineskip}{\uppercase}
\titleformat*{\section}{\bfseries\fontsize{12pt}{12pt}}
\titleformat*{\subsection}{\bfseries\fontsize{12pt}{12pt}}

% Chapter title spacing
\titlespacing{\chapter}{0pt}{25mm}{3\baselineskip}

% EPIGRAPH
\setlength{\epigraphwidth}{.6\textwidth}
\setlength{\epigraphrule}{0pt}

\hypersetup{
    pdftitle			= {MFA Thesis},
    pdfauthor			= {Mustafa ilhan},
    pdfsubject			= {Trash, Art},
    pdfkeywords			= {mfa thesis, trash, art},
    colorlinks			= false,
    pdfborder			= {0 0 0},
    pdfpagemode			= UseOutlines
}

\providecommand{\quotes}[1]{``#1''}
\providecommand{\singlequotes}[1]{`#1'}
\providecommand{\comment}[1]{\textcolor{cyan}{#1}}
\providecommand{\summary}[1]{\textcolor{blue}{[#1]}}
\providecommand{\paraphrase}[1]{\textcolor{magenta}{\quotes{#1}}}

% RENAMING CHAPTER TITLES
\addto\captionsenglish{%
  \renewcommand{\contentsname}{\MakeUppercase{Table of Contents}}%
  \renewcommand{\listtablename}{\MakeUppercase{List of Tables}}%
  \renewcommand{\listfigurename}{\MakeUppercase{List of Figures}}%
  \renewcommand{\bibname}{\MakeUppercase{Bibliography}}%
}

%****************************************
% Title Ideas for Thesis 
% - Trash as an Artistic Medium
% - Notebooks from Trash: Trash in the Context of Art
% - Notebooks from Trash: Exploring the potential of trash through art practice
% - Transforming Trash as an Artistic Act
% - Act of Transforming Trash in the Context of Art
%........................................

%****************************************
% BEGIN
\begin{document}
\pagenumbering{gobble}
\selectlanguage{english}

%****************************************
% COVER
\begin{titlepage}
    \begin{center}

        \singlespacing

        \vspace*{25mm}
        \uppercase{Transforming Trash as an Artistic Act}

        \vspace{15mm}
        A Master’s Thesis\\
        \vspace{15mm}
        by\\
        MUSTAFA İLHAN

        \vfill
        Department of\\
        Communication and Design\\
        İhsan Doğramacı Bilkent University\\
        Ankara\\
        January 2016
        \vspace{25mm}

    \end{center}
\end{titlepage}
%........................................

%****************************************
% BLANK PAGE
\clearpage
\afterpage{\null\newpage}
\clearpage
%........................................

%****************************************
% DEDICATION
\newenvironment{dedication}
  {\clearpage           % we want a new page
   \thispagestyle{empty}% no header and footer
   \vspace*{2in}		% some space at the top 
   \centering
  }
  {\par					% end the paragraph
   \vfill
   \clearpage           % finish off the page
  }

\begin{dedication}
To my family and for those who embrace trash
\end{dedication}
%........................................

\pagenumbering{roman}

%****************************************
% TITLE PAGE
\clearpage
\thispagestyle{empty}
\singlespacing
	\begin{center}

        \vspace*{25mm}
        TRANSFORMING TRASH AS AN ARTISTIC ACT\\
        \vspace*{4\baselineskip}
        Graduate School of Economics and Social Sciences\\
        of\\
		İhsan Doğramacı Bilkent University\\
		\vspace*{3\baselineskip}
        by\\
       	\vspace*{3\baselineskip}
        MUSTAFA İLHAN\\
        \vspace*{4\baselineskip}
        In Partial Fulfilment of the Requirements for the Degree of\\
		MASTER FINE OF ARTS\\
        \vspace{2\baselineskip}
        in\\
        \vspace{2\baselineskip}
        THE DEPARTMENT OF\\
        COMMUNICATION AND DESIGN\\
       	İHSAN DOĞRAMACI BİLKENT UNIVERSITY\\
        ANKARA\\
        \vspace{\baselineskip}
        January 2016
	\end{center}
\clearpage
%........................................

%****************************************
% APPROVAL PAGE
\clearpage
\thispagestyle{empty}
\singlespacing
\noindent I certify that I have read this thesis and have found that it is fully adequate, in scope and in quality, as a thesis for the degree of Master of Fine Arts in Media and Design.\\

\noindent---------------------------------\\
Assist. Prof. Dr. Ersan Ocak\\
Supervisor\\

\noindent I certify that I have read this thesis and have found that it is fully adequate, in scope and in quality, as a thesis for the degree of Master of Fine Arts in Media and Design.\\

\noindent---------------------------------\\
Prof. Dr. Mehmet Yılmaz\\
Examining Commitee Member\\

\noindent I certify that I have read this thesis and have found that it is fully adequate, in scope and in quality, as a thesis for the degree of Master of Fine Arts in Media and Design.\\

\noindent---------------------------------\\
Instructor Ekin Kılıç\\
Examining Commitee Member\\

\noindent Approval of the Graduate School of Economics and Social Sciences\\

\noindent---------------------------------\\
Prof. Dr. Halime Demirkan\\
Director\\

\clearpage
%........................................

%****************************************
% ENGLISH ABSTRACT
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{ABSTRACT}
\doublespacing
\begin{center}
	\vspace*{25mm}
	\MakeUppercase{Abstract}\\
    \vspace{\baselineskip}
    TRANSFORMING TRASH AS AN ARTISTIC ACT\\
    İlhan, Mustafa\\
    M.F.A., Department of Communication and Design\\
    Supervisor: Assist. Prof. Dr. Ersan Ocak\\
    \vspace{\baselineskip}
    January 2016
\end{center}
\singlespacing
\par Usage of non-art objects in the artworks can be traced back to the beginning of 20th century. This approach has broken the existing rules and extended the language of art in an unprecedented way. After that point, trash has found a place in the work of artists \comment{(and their expression)}.

The purpose of this study is to discuss transformation (or usage) of trash in the context of (art and) artistic act. It is questioned that the role of art in the transformation of discarded items. Approaches and methodologies of artists who use discarded materials in their works are explored. This thesis follows the discussions of the place of trash in culture and theories of trash. Then approaches and methodologies of artists who use discarded materials in their works are explored. \comment{Usage of discarded materials in the artwork by the contemporary artist are studied through their works.} It is questioned that why and how they use discarded materials and how artworks can be understood regarding theoretical and cultural aspects.

Within this framework, thesis project argues that it is possible to see the process of collecting, transforming and placing trash as an artistic act. In the scope of the project, discarded papers are collected with the help people and transformed to notebooks. Progress and collected items are recorded on a website. Notebooks are installed in different locations in order to invite people to use trash again.

\noindent\textbf{Keywords:} Transformation of Trash, Trash in Art, Trash as an Art Material.
\clearpage
%........................................

%****************************************
% TURKISH ABSTRACT
\thispagestyle{plain}
\selectlanguage{turkish}
\addcontentsline{toc}{chapter}{ÖZET}
\doublespacing
\begin{center}
	\vspace*{25mm}
	\MakeUppercase{Özet}\\
    \vspace{\baselineskip}
    \MakeUppercase{Sanatsal Bir Eylem Olarak Çöpü Dönüştürme}\\
    İlhan, Mustafa\\
    Yüksek Lisans, İletişim ve Tasarım Bölümü\\
    Tez Yöneticisi: Yrd. Doç. Dr. Ersan Ocak\\
    \vspace{\baselineskip}
    Ocak 2016
\end{center}
\singlespacing
\par \comment{Sanatta çöpün yeri incelenip, çöpü dönüştürmenin bir sanatsal eylem olduğu iddia edilmektedir. Ya da en azından proje kapsamında üretilen işin. Ya da bu bağlamda ele alınabileceği iddia edilmektedir.}

Bu çalışmanın amacı çöpü dönüştürmenin sanat ve sanatsal eylem bağlamında tartışmaktır. Çöpün kültürel ve kuramsal olarak nasıl ele alındığı(alınabileceği) araştırılmıştır. Çöpü dönüştürmenin nasıl ve ne şekilde mümkün olduğu incelenmiştir. Sanatsal üretim yöntemi olarak çöpü (tercih eden)/(kullanan) sanatçıların yaklaşımları ele alınmıştır. Neden ve nasıl atıkları sanat işlerinde ve eylemlerinde kullandıkları sorgulanmıştır ve bunun toplumsal ve teorik olarak nereye oturduğu/ne anlama geldiği/nasıl anlaşılabileceği incelenmiştir.

\comment{Tez dahilinde geliştirilen proje ile çöpten dönüştürülerek üretilen defterler ile çöpe bakış açısının dönüştürülmesi amaçlanmıştır. Farklı zaman ve mekanlardan toplanan kağıtlar birbirinden farklı kompozisyonlar oluşturmakta ve endüstriyel defterlere alternatif oluşturmaktadır. Bu defter hakim sistemin dışındaki mekanlarda konulmuştur. (kamusal ortak alan)}

\comment{Sanat dışı objelerin sanata malzeme olması sanatın dilinde genişlemere olanak vermiştir (ve çöpün sanat işlerinden kullanılması bu bağlamda incelenebilmektedir.). Beraberinde pek çok akım ve üretim biçimini tetiklemiştir. Hem sanatın dili genişledi hem de objeler sanat için farklı anlamlarda ve amaçlarla kullanılmaya başladı. Bu çalışmda çöpün sanatsal bağlamda ele alınması yapılacaktır. Çöpün nasıl sanatta yer bulduğu incelenmiştir.}

Çöp ve sanatsal eylem birlikte alınarak incelenmiştir. Çöp olgusu ve sanat dışı malzemelerin sanatta kullanılması sorgulanmıştır. Teoriler ve methodlar ele alınmıştır. Çöpün ne olduğu sorgulanmaktadır ve çöpe yeni bir bakış açısı kazandırmak amaçlanmaktadır.

Tez dahilinde geliştirilen proje için insanlar örgütlenerek farklı mekanlardan farklı tiplerde atık kağıtlar toplanmıştır. Bu atık maddeler dönüştürülerek defter yapılmıştır. Renk, doku, biçim ve üretim yöntemi ile endüstriyel defterlere alternatif bir yaklaşım olması amaçlanmıştır. Bu defterler insanların ulaşabileceği belirli mekanlara dağıtılmıştır (yerleştirilmiştir). Mekanlar da aynı şekilde eleştirel bakış açısına açık yerler arasından seçilmiştir. Alternatif bir ortam sunan mekanlar seçilmiştir. Defterlerin mekanla belirli bir anlamsal ilişki içine girmesi amaçlanmıştır. Bu proje sonunda dönüştürülen şeylerin aynı zamanda dönüşütüğü şeyi de başka bir şeye dönüştürdüğü anlaşılmaktadır. Çöpe olan bakış açısının dönüştürülmesi amaç edilmiştir.

\noindent\textbf{Anahtar Kelimeler:} Çöpün Dönüştürülmesi, Sanatta Çöp.
\clearpage
%........................................

\selectlanguage{english}

%****************************************
% ACKNOWLEDGMENTS
\thispagestyle{plain}
\addcontentsline{toc}{chapter}{ACKNOWLEDGMENTS}
\doublespacing
\begin{center}
	\vspace*{25mm}
	\MakeUppercase{ACKNOWLEDGMENTS}\\
    \vspace{3\baselineskip}
\end{center}
\par First and foremost, I would like to thank to my thesis supervisor Ersan Ocak for his continuous patience, support, encouragement and immense knowledge that he shared with me through the entire process. I cannot imagine having a better advisor for both this written thesis and the project I conducted. I also would like to thank the rest of my thesis committee; Mehmet Yılmaz and Ekin Kılıç for their motivation and insightful comments.
\clearpage
%........................................

%****************************************
% TABLE OF CONTENTS
\addcontentsline{toc}{chapter}{Table of Contents}

\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

\renewcommand{\cftchapleader}{\cftdotfill{\cftdotsep}} % dots for chapters
\renewcommand{\cfttoctitlefont}{\normalfont\MakeUppercase\hfil}
\renewcommand\cftchapfont{\mdseries}
\renewcommand\cftchappagefont{\mdseries}
\renewcommand{\cftchappresnum}{Chapter\space}
\renewcommand{\cftchapaftersnum}{:}

\setlength{\cftbeforetoctitleskip}{25mm}
\setlength{\cftaftertoctitleskip}{3\baselineskip}
\setlength{\cftchapnumwidth}{5em}
%\addtolength{\cftchapnumwidth}{\cftchappresnum\cftchapaftersnum}

\begin{singlespace}
\tableofcontents
\end{singlespace}
\clearpage
%........................................

%****************************************
% LIST OF FIGURES
\addcontentsline{toc}{chapter}{List of Figures}

\renewcommand{\cftloftitlefont}{\normalfont\MakeUppercase\hfil}
\renewcommand{\cftfigaftersnum}{.}

\setlength{\cftbeforeloftitleskip}{25mm}
\setlength{\cftafterloftitleskip}{3\baselineskip}

\begin{singlespace}
\listoffigures
\end{singlespace}
\clearpage
%{%
%\let\oldnumberline\numberline%
%\renewcommand{\numberline}{\figurename~\oldnumberline}%
%\listoffigures%
%}
%........................................

%****************************************
% LIST OF FIGURES
%\addcontentsline{toc}{chapter}{List of Tables}
%\listoftables
%........................................

%\listoftodos

\pagenumbering{arabic}

%****************************************
% CHAPTERS
\include{chapter1_reduced}
\include{chapter2_reduced}
\include{chapter3_reduced}
\include{chapter4_reduced}
\include{chapter5}
%........................................

%****************************************
% BIBLIOGRAPHY
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{apacite}
\begin{singlespace}
\setlength{\bibsep}{2\itemsep}
\bibliography{literature}
\end{singlespace}
\clearpage
%........................................

%****************************************
% APENDICES
\begin{appendices}
\include{appendix_glossary}
\end{appendices}
%........................................

\end{document}