### 2015-12-29 appendix_materials.tex
Committee Copy
Former-commit-id: 4798a7cb56c31623135a8d42eb713cb8e932ebe7

\appendix
\chapter{Collected materials}

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.47\textwidth}
        \includegraphics[width=\textwidth]{project_graphics/collected1.jpg}
    \end{subfigure}
    \begin{subfigure}[t]{0.47\textwidth}
        \includegraphics[width=\textwidth]{project_graphics/collected2.jpg}
    \end{subfigure}
    \label{fig:CollectedMaterials1}
\end{figure}