### 2016-01-17 main.tex

\documentclass[12pt]{report}
\usepackage[hidelinks]{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[turkish,english,shorthands=:!]{babel}
\usepackage{apacite}
% left: 40mm, top: 25mm, right: 25mm, bottom: 25mm
\usepackage[a4paper,left=40.2mm,top=24.2mm,right=25.4mm,bottom=36mm]{geometry}
\usepackage{epigraph}
\usepackage{graphicx}
\usepackage{float}
\usepackage[font={footnotesize,bf}]{caption}
\usepackage[font={footnotesize,bf}]{subcaption}
\usepackage{wrapfig}
\usepackage[usenames,dvipsnames]{color}
\usepackage[titletoc,title]{appendix}
\usepackage{afterpage}
\usepackage[doublespacing]{setspace}
\usepackage{tocloft}
\usepackage{url}
\usepackage{indentfirst}

% Turning off all figures, required for turnitin upload
% Remove the next two lines if you want the figures at their place
%\usepackage[figuresonly,nolists,nomarkers]{endfloat}
%\renewcommand{\processdelayedfloats}{}

% Same fonts for URL
\AtBeginDocument{\urlstyle{APACsame}}

% Figure and Table counters
\usepackage{chngcntr}
\counterwithout{figure}{chapter}
\counterwithout{table}{chapter}

%****************************************
% SPACING

% Footnote line spacing
\setlength{\footnotesep}{\baselineskip}

% Adding extra space between paragraphs
\setlength{\parskip}{0.5\baselineskip}
%........................................

%****************************************
% BIBLIOGRAPHY

% Bibliography line spacing
\usepackage{natbib}
\setlength{\bibsep}{0.0pt}

% Removing parentheses around year in bibliography
\AtBeginDocument{
\renewcommand{\BBOP}{}
\renewcommand{\BBCP}{}
}

% Adding ":" character before page number
\bibpunct[: ]{(}{)}{:}{a}{,}{~}
%........................................

%****************************************
% BLOCK QUOTES

% Spaces before and after block quotes
\usepackage{etoolbox}
\AtBeginEnvironment{quote}{\vspace{-\baselineskip}}
\AtEndEnvironment{quote}{\vspace{-\baselineskip}}

% Spaces before and after block quotes, indentation
\renewcommand{\quote}{\list{}{\rightmargin=4em\leftmargin=4em}\item\relax}

% Spacing
\expandafter\def\expandafter\quote\expandafter{\quote\singlespacing}
%........................................

%****************************************
% CHAPTERS

% Chapter titles
\usepackage{titlesec}
\titleformat{\chapter}[display]
        {\normalfont\fontsize{14pt}{14pt}\centering\bfseries}{\MakeUppercase{\chaptertitlename}\  \thechapter}{2\baselineskip}{\uppercase}
\titleformat*{\section}{\bfseries\fontsize{12pt}{12pt}}
\titleformat*{\subsection}{\bfseries\fontsize{12pt}{12pt}}

% Chapter title spacing
% Centered, Top: 50mm, Next: 3 line after
\titlespacing{\chapter}{0pt}{13mm}{3\baselineskip}

% Bibliography title spacing
% Centered, Top: 50mm, Next: 4 line after
\titlespacing*{name=\chapter,numberless}{0pt}{16mm}{4\baselineskip}

% RENAMING CHAPTER TITLES
\addto\captionsenglish{%
  \renewcommand{\contentsname}{\MakeUppercase{Table of Contents}}%
  \renewcommand{\listtablename}{\MakeUppercase{List of Tables}}%
  \renewcommand{\listfigurename}{\MakeUppercase{List of Figures}}%
  \renewcommand{\bibname}{\MakeUppercase{Bibliography}}%
}
%........................................

%****************************************
% EPIGRAPH

% Epigraph setup
\setlength{\epigraphwidth}{.65\textwidth}
\setlength{\epigraphrule}{0pt}
%........................................

% PDF setup
\hypersetup{
    pdftitle	= {MFA Thesis},
    pdfauthor	= {Mustafa ilhan},
    pdfsubject	= {Trash, Art, Transformation},
    pdfkeywords	= {mfa thesis, trash, art, transformation},
    colorlinks	= false,
    pdfborder	= {0 0 0},
    pdfpagemode	= UseOutlines
}

% Utilities
\providecommand{\quotes}[1]{``#1''}
\providecommand{\singlequotes}[1]{`#1'}
\providecommand{\comment}[1]{\textcolor{cyan}{#1}}
\providecommand{\summary}[1]{\textcolor{blue}{[#1]}}
\providecommand{\paraphrase}[1]{\textcolor{magenta}{\quotes{#1}}}

%****************************************
% BEGIN
\begin{document}
\pagenumbering{gobble}
\selectlanguage{english}

%****************************************
% COVER
\begin{titlepage}
\singlespacing
    \begin{center}

        \vspace*{20mm}
        \uppercase{Transforming Trash as an Artistic Act}\\
        \vspace*{3\baselineskip}
        A Master’s Thesis\\
        \vspace*{3\baselineskip}
        by\\
        MUSTAFA İLHAN

        \vfill
        Department of\\
        Communication and Design\\
        İhsan Doğramacı Bilkent University\\
        Ankara\\
        January 2016
        \vspace*{28mm}

    \end{center}
\end{titlepage}
%........................................

%****************************************
% BLANK PAGE
\clearpage
\afterpage{\null\newpage}
\clearpage
%........................................

%****************************************
% DEDICATION
\newenvironment{dedication}
  {\clearpage           % we want a new page
   \thispagestyle{empty}% no header and footer
   \vspace*{2in}		% some space at the top 
   \centering
  }
  {\par					% end the paragraph
   \vfill
   \clearpage           % finish off the page
  }

\begin{dedication}
To my family and for those who embrace trash
\end{dedication}
%........................................

\pagenumbering{roman}

%****************************************
% TITLE PAGE
\clearpage
\thispagestyle{empty}
\singlespacing
	\begin{center}

        \vspace*{25mm}
        TRANSFORMING TRASH AS AN ARTISTIC ACT\\
        \vspace*{4\baselineskip}
        Graduate School of Economics and Social Sciences\\
        of\\
		İhsan Doğramacı Bilkent University\\
		\vspace*{2\baselineskip}
        by\\
       	\vspace*{2\baselineskip}
        MUSTAFA İLHAN\\
        \vspace*{3\baselineskip}
        In Partial Fulfilment of the Requirements for the Degree of\\
		MASTER OF FINE ARTS\\
        \vspace{\baselineskip}
        in\\
        \vspace{\baselineskip}
        THE DEPARTMENT OF\\
        COMMUNICATION AND DESIGN\\
       	İHSAN DOĞRAMACI BİLKENT UNIVERSITY\\
        ANKARA\\
        \vspace{\baselineskip}
        January 2016

	\end{center}
\clearpage
%........................................

%****************************************
% APPROVAL PAGE
\clearpage
\thispagestyle{empty}
%\singlespacing
%\doublespacing
\noindent I certify that I have read this thesis and have found that it is fully adequate, in scope and in quality, as a thesis for the degree of Master of Fine Arts in Media and Design.\\

\noindent---------------------------------\\
Assist. Prof. Dr. Ersan Ocak\\
Supervisor\\

\noindent I certify that I have read this thesis and have found that it is fully adequate, in scope and in quality, as a thesis for the degree of Master of Fine Arts in Media and Design.\\

\noindent---------------------------------\\
Prof. Dr. Mehmet Yılmaz\\
Examining Committee Member\\

\noindent I certify that I have read this thesis and have found that it is fully adequate, in scope and in quality, as a thesis for the degree of Master of Fine Arts in Media and Design.\\

\noindent---------------------------------\\
Instructor Ekin Kılıç\\
Examining Committee Member\\

\noindent Approval of the Graduate School of Economics and Social Sciences\\

\noindent---------------------------------\\
Prof. Dr. Halime Demirkan\\
Director\\

\clearpage
%........................................

%****************************************
% ENGLISH ABSTRACT
\thispagestyle{plain}
\phantomsection
\addcontentsline{toc}{chapter}{ABSTRACT}
\doublespacing
\begin{center}
	\vspace*{13mm} % Top: 50mm
	{\fontsize{14pt}{14pt}\selectfont \textbf{\MakeUppercase{Abstract}}}\\
    \vspace{\baselineskip}
    TRANSFORMING TRASH AS AN ARTISTIC ACT\\
    İlhan, Mustafa\\
    M.F.A., Department of Communication and Design\\
    Supervisor: Assist. Prof. Dr. Ersan Ocak\\
    \vspace{\baselineskip}
    January 2016
\end{center}
%\singlespacing
\par Usage of non-art objects in the artworks can be traced back to the beginning of twentieth century. This approach has broken the existing rules and extended the language of art in an unprecedented way. After that point, trash has found a place in the work of artists.

The purpose of this study is to discuss usage of trash in the context of art. This thesis questions the role of art in the transformation of discarded items through the approaches and methods of artists who use discarded materials in their works. It examines why and how artists use discarded materials and how their artworks can be understood regarding theoretical and cultural aspects. This study is supported by theoretical and philosophical ideas about garbage.

Within this framework, thesis project argues that it is possible to see the process of transforming trash as an artistic act. In the scope of the project, discarded papers are collected with the help people and transformed to notebooks. Progress and collected materials are recorded on a website. Notebooks are installed in non-conventional art spaces such as library desks and computer labs that are frequently visited common places in order to invite people to use trash again.

\noindent Keywords: Transformation of Trash, Trash in Art, Trash as an Art Material.
\clearpage
%........................................

%****************************************
% TURKISH ABSTRACT
\thispagestyle{plain}
\selectlanguage{turkish}
\phantomsection
\addcontentsline{toc}{chapter}{ÖZET}
\doublespacing
\begin{center}
	\vspace*{13.5mm} % Top: 50mm
	{\fontsize{14pt}{14pt}\selectfont \textbf{\MakeUppercase{Özet}}}\\
    \vspace{\baselineskip}
    \MakeUppercase{Sanatsal Bir Eylem Olarak Atığı Dönüştürme}\\
    İlhan, Mustafa\\
    Yüksek Lisans, İletişim ve Tasarım Bölümü\\
    Tez Yöneticisi: Yrd. Doç. Dr. Ersan Ocak\\
    \vspace{\baselineskip}
    Ocak 2016
\end{center}
%\singlespacing
\par Sanat işlerinde sanat dışı nesnenin kullanımı yirminci yüzyılın başına kadar takip edilebilmektedir. Bu yaklaşım var olan kuralları yıkmış ve sanatın dilini eşi görülmemiş bir biçimde genişletmiştir. Bu noktadan sonra atık sanatçıların işlerinde yer bulmuştur.

Bu çalışmanın amacı sanatsal bağlamda atığın kullanımını tartışmaktır. Atık ürünlerin dönüştürülmesinde sanatın rolü, işlerinde atık maddeleri kullanan sanatçıların yaklaşımları ve yöntemleri üzerinden sorgulanmaktadır. Sanatçıların neden ve nasıl atık malzemeleri kullandıkları ve işlerinin kültürel ve kuramsal açılardan nasıl anlaşılabileceği araştırılmaktadır. Bu çalışma çöpe dair kuramsal ve felsefi fikirler ile desteklenmektedir.

Bu çerçevede, tez projesi atığı dönüştürme sürecinin bir sanatsal eylem olarak görülebileceğini savunmaktadır. Proje kapsamında atık kağıtlar insanların yardımıyla toplanıp not defterlerine dönüştürülmüştür. Süreç ve toplanan malzemeler bir web sitesinde kayıt altına alınmıştır. İnsanları atığı tekrar kullanmaya davet etmek amacıyla, defterler geleneksel sanat mekanı olarak kabul edilmeyen kütüphane, bilgisayar laboratuvarı gibi sık ziyaret edilen ortak kullanım alanlarına yerleştirilmiştir.

\noindent Anahtar Kelimeler: Atığın Dönüştürülmesi, Sanatta Atık, Sanatsal Malzeme olarak Atık.
\clearpage
%........................................

\selectlanguage{english}

%****************************************
% ACKNOWLEDGMENTS
\thispagestyle{plain}
\phantomsection
\addcontentsline{toc}{chapter}{ACKNOWLEDGMENTS}
\doublespacing
\begin{center}
	\vspace*{13mm} % Top: 50mm
	{\fontsize{14pt}{14pt}\selectfont \textbf{\MakeUppercase{ACKNOWLEDGMENTS}}}\\
    \vspace{3\baselineskip}
\end{center}
\par First and foremost, I would like to thank to my thesis supervisor Assist. Prof. Dr. Ersan Ocak for his endless patience, support and encouragement through the whole progress. He contributed in every way he can and guided me through the development of this thesis with his immense knowledge. I also would like to thank the rest of my thesis committee; Prof. Dr. Mehmet Yılmaz and Instructor Ekin Kılıç for their valuable contributions and insightful comments on the issues of this thesis.

I am grateful for Assist. Prof. Dr. Marek Brzozowski, Assist. Prof. Dr. Andreas Treske and Assist. Prof. Dr. Ahmet Gürata for their valuable ideas and criticisms.

I would like to thank my colleagues; Mert Aslan and Eda Erdem for their friendship and comments.

I must thank to Melis Golar and Cemil Batur Gökçeer from Torun for opening their space to my project.

I would like to thank my bosses and colleagues from Mobilus; Özgür Deniz Önür, Yağız Yaşaroğlu, Ekin Dino, Ayşegül Yaşaroğlu, Süleyman Yıldız and Aytaç Kocaman for their support, trust and collecting their discard for the thesis project.

I am grateful to M. Elgin Akpınar, Nihan Özdemirler, M. Cem B. Üstündağ, B. Ceren D. Üstündağ, Ayşenur Durgut, Emrah Şamdan, Duygu Ç. Şamdan, and Koray Benli for their valuable friendship, and also Mr. Ali from METU Technopolis İş Bank Office, Mr. Aykut from Modshifters, and Varuna Gezgin for collecting discarded materials for the thesis project.

Lastly and mostly, I would like to thank my family for their love, trust and support from the very beginning of this study. 
\clearpage
%........................................

%****************************************
% TABLE OF CONTENTS
\phantomsection
\addcontentsline{toc}{chapter}{TABLE OF CONTENTS}

\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

\renewcommand{\cftchapleader}{\cftdotfill{\cftdotsep}} % dots for chapters
\renewcommand{\cfttoctitlefont}{\MakeUppercase\hfil\bfseries\fontsize{14pt}{14pt}\selectfont}
\renewcommand\cftchapfont{\mdseries}
\renewcommand\cftchappagefont{\mdseries}
\renewcommand{\cftchappresnum}{CHAPTER\space}
\renewcommand{\cftchapaftersnum}{:}

\setlength{\cftbeforetoctitleskip}{20mm}
\setlength{\cftaftertoctitleskip}{3\baselineskip}
\setlength{\cftchapnumwidth}{7em}
%\addtolength{\cftchapnumwidth}{\cftchappresnum\cftchapaftersnum}

% Spacing between items
\setlength{\cftbeforesecskip}{\cftbeforechapskip}
\setlength{\cftbeforesubsecskip}{\cftbeforechapskip}

\begin{singlespace}
\tableofcontents
\end{singlespace}
\clearpage
%........................................

%****************************************
% LIST OF FIGURES
\phantomsection
\addcontentsline{toc}{chapter}{LIST OF FIGURES}

% Spacing between items
\makeatletter
\def\@chapter[#1]#2{\ifnum \c@secnumdepth >\m@ne
	\refstepcounter{chapter}%
	\typeout{\@chapapp\space\thechapter.}%
	\addcontentsline{toc}{chapter}%
	{\protect\numberline{\thechapter}#1}%
	\else
	\addcontentsline{toc}{chapter}{#1}%
	\fi
	\chaptermark{#1}%
	\if@twocolumn
	\@topnewpage[\@makechapterhead{#2}]%
	\else
	\@makechapterhead{#2}%
	\@afterheading
	\fi}
\makeatother

\renewcommand{\cftloftitlefont}{\MakeUppercase\hfil\bfseries\fontsize{14pt}{14pt}\selectfont}
\renewcommand{\cftfigaftersnum}{.}

\setlength{\cftbeforeloftitleskip}{20mm}
\setlength{\cftafterloftitleskip}{3\baselineskip}
\setlength{\cftbeforefigskip}{\cftbeforechapskip}

\begin{singlespace}
\listoffigures
\end{singlespace}
\clearpage
%........................................

%****************************************
% LIST OF TABLES
%\phantomsection
%\addcontentsline{toc}{chapter}{LIST OF TABLES}
%\listoftables
%........................................

\pagenumbering{arabic}

%****************************************
% CHAPTERS
\include{chapter1_final}
\include{chapter2_final}
\include{chapter3_final}
\include{chapter4_final}
\include{chapter5_final}
%........................................

%****************************************
% BIBLIOGRAPHY
\nocite{*} % Add all references
\phantomsection
\addcontentsline{toc}{chapter}{BIBLIOGRAPHY}
\bibliographystyle{apacite}
\begin{singlespace}
\setlength{\bibsep}{2\itemsep}
\bibliography{literature}
\end{singlespace}
\clearpage
%........................................

%****************************************
% APENDICES
\appendix
\phantomsection
\addtocontents{toc}{\cftpagenumbersoff{chapter}} 
\chapter*{APPENDICES}
\addtocontents{toc}{\protect\contentsline {chapter}{APPENDICES}{}{}}

\addtocontents{toc}{\cftpagenumberson{chapter}}

\makeatletter
\addtocontents{toc}{\let\protect\l@chapter\protect\l@section}
\makeatother

\include{appendix_materials}
\include{appendix_project}
%........................................

\end{document}